---
title: "Gov 50 Final Project"
author: Sarah Park
description: "My final project"
output:
  distill::distill_article:
    self_contained: false
---



## Project thoughts

I am interested in exploring data related to how the environmental index in different countries relates to the mortality rate in these different countries. I think this is interesting because environmental harm such as pollution can impact mortality rate as it can cause health issues.

##Milestone 2
I will use data on billionaires from Kaggle: https://www.kaggle.com/datasets/nelgiriyewithana/billionaires-statistics-dataset/. My research question will be: is there a correlation between self made status and final net worth? My variables of interest are whether an individual is self made or if they have inherited wealth, an their final net worth. My hypothesis is that self made billionaires are more likely to have higher net worths as they likely create something innovative while generational wealth may be lost over time. The variation is the difference in status of self made or not self made. An observed pattern in the data that would provide support for my hypothesis would be that self made billionaires would have a higher net worth as they create new companies and innovations while generational wealth may be diluted over time. My sample is comprised of data from Kaggle from 2023 which contains statistics on the world's billionaires and other information. My outcome variable would be net worth.


##Milestone 3
```{r}
library(readr)
library(ggplot2)
library(tidyverse)
library(dplyr)
mydata <- read_csv("~/Desktop/Billionaires Statistics Dataset.csv", show_col_types = FALSE)
mydata$selfMade <- as.integer(as.logical(mydata$selfMade))
mydata
```


```{r}
mydata <- mydata |>
 mutate(self_made = if_else(selfMade == 1, "Self made", "Not self made")) 

my_data_scatter <- mydata |>
  ggplot(mapping = aes(x = self_made, y = finalWorth, color = self_made)) + geom_point() +
  labs(
    x = "Self Made Status",
    y = "Net Worth (in 10000 USD)",
    color = "Self Made Status")

my_data_scatter
```





